\chapter{Introduction}

\section{Motivation}

Options are derivative contracts that give the holder the right, but not the obligation, to buy or sell an underlying asset at a specified strike price $K$, specified at the time of option purchase. The act of buying or selling the underlying asset is called exercising the option, and is done at a predetermined expiry time $T$ (for European options) or at any time before expiration (for American options). A contract allowing the holder to buy the underlying asset is called a call option, and that allowing the holder to sell the underlying asset is called a put option.

These options are very important financial instruments, and tens of millions of contracts are traded every day just for the S\&P 500 index alone, as they are useful for hedging and speculation. For that reason, it is important to be able to price these options accurately. This is a non-trivial task, as the price of an option depends on many factors, such as the underlying asset price which is generally modelled as a stochastic process.

\section{Literature review}

The pioneering work in option pricing was done by Black and Scholes \cite{black1973pricing}, who derived a closed-form solution for the price of the European call option under the assumption that the underlying asset price follows a geometric Brownian motion. This model, known as the Black-Scholes model, is considered the foundational work in option pricing theory. However, it relies on several assumptions such as constant risk-free rate and volatility, which is not always realistic in practice. As a result, Heston \cite{heston1993closed} derived the European option prices of the proposed stochastic volatility model, allowing the volatility to vary over time to capture more facets of market dynamics.

The American option is much more difficult to price, since it allows the holder to exercise the option at any time before expiration, making it analogous to an optimal stopping problem that does not admit a closed-form solution in general. As a result, various numerical methods have been developed to tackle this problem, such as the binomial tree method \cite{cox1979option, vellekoop2009tree} for Black-Scholes and Heston respectively, Monte-Carlo simulation methods \cite{glasserman2004monte} and more recently variational methods where the optimal stopping problem is formulated as a variational inequality, solved by deep neural networks \cite{raissi2019physics, bai2022application}. This method is adapted to incoporate fractional Sobolev norms to guarantee convergence of the solution to the true option price instead of just pointwise convergence, which has been particularly developed for American option pricing \cite{zhao2025fractional}.

\section{Outline of project}

Our project focuses on the Black-Scholes and Heston models, for which we will progressively build benchmarks for increasingly complex options. For each model, we begin by deriving the price of a European call option, which serves as a benchmark for the binomial tree method. The binomial tree easily extends to American options, providing a benchmark for the variational neural network method, which we implement with both the standard loss function and the Sobolev loss function. We then extend the analysis to multi-asset options. We first price options on the product of assets, since there exists an equivalent single-asset reduction and can be validated against the benchmarks already established. This foundation allows us to tackle payoffs without such a reduction, such as the maximum or minimum of several assets. Once the numerical results across these methods agree, we will apply our pricing scheme to two further objectives: computing the free boundary of the American option under varying market conditions, and evaluating the Greeks (the sensitivities of the option price to model parameters) to gain insight into exposures and hedging strategies.
